Upwelling variability across the Benguela upwelling systems

### 2019-09-24 SST_patterns.Rmd
update

```{r}
# Generalosed linear models
#Using the generalised linear hypotheses - Look at the difference in metrics (duration and intensity) of in situ upwelling and remotely sensed SST data upwelling

@@

@@

```{r}
# Which pixels showed the highest upwelling counts within the different datasets
# Which site showed the most intense upwelling within the different datasets
# Where is the largest upwelling event occuring?

@@

@@

- Pvalues
- plotting
- ERA5
- Plot wind data

>>>>

@@

```{r}
# Generalosed linear models

# Generalized Linear Models- Look at the difference in metrics of in situ upwelling and remotely sensed SST data upwelling 
# Generalized linear models are fit using the glm( ) function. The form of the glm function is glm(formula, family=familytype(link=linkfunction), data=)

trends_duration_func <- function(df){
  upwellmodel <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$duration[SACTN_upwell_base$event_no])), 
                     family = poisson(link = "log"))
  upwellmodel0 <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ 1, family = poisson(link = "log")) # intercept only
  dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
                       upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's pseudo-R2
                       upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
}

OISST_trends_duration_func <- trends_duration_func(df = OISST_upwell_base)
CMC_trends_duration_func <- trends_duration_func(df = CMC_upwell_base)
SACTN_trends_duration_func <- trends_duration_func(df = SACTN_upwell_base)

### intensoty 
trends_intensity_func <- function(df){
  upwellmodel <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no])), 
                     family = gaussian(link = "identity"))
  upwellmodel0 <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ 1, family = gaussian(link = "identity")) # intercept only
  dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
                       upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's psesudo-R2
                       upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
}

OISST_trends_intensity_func <- trends_intensity_func(df = OISST_upwell_base)
CMC_trends_intensity_func <- trends_intensity_func(df = CMC_upwell_base)
SACTN_trends_intensity_func <- trends_intensity_func(df = SACTN_upwell_base)
```

@@

```{r}
# Which pixels showed the highest upwelling counts within the different datasets
# Which site showed the most intense upwelling within the different datasets

@@

# Wind plots

```{r, fig.cap = "Windrose diagram representing the wind direction and speed for each of the sites", fig.height = 10, fig.width = 20}
library(ggradar)
library(dplyr)
library(scales)
library(tibble)
load("Data/UI_angle.RData")

# Detects NA values, no NA values are present
# Plotting wind variables
source("Functions/wind.rose.R") # This it may be related to the countmax = NA in the function

wind_daily_renamed <- UI_angle %>% 
  dplyr::rename(spd = mean_speed) %>%
  dplyr::rename(dir = dir_circ)

p.wr2 <- plot.windrose(data = wind_daily_renamed,
              spd = "spd",
              dir = "dir")

p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 4, nrow = 5) +
  theme(strip.text.x = element_text(size = 25))
p.wr3

################3

load("Data/UI_angle.RData")

wind_plot_prep <- UI_angle %>% 
  select(-t, -ui.saws)

wind_plot <- wind_plot_prep %>% 
  as_tibble(rownames = "site") %>% 
  group_by(site) %>% 
  mutate_at(vars(-site), rescale) %>% 
  tail(4)

ggradar(wind_plot)

??countmax

# ANOVA
- Relationship between duration, year and distance as a function of the mean intensity.

```{r}
anova_func <- function(df){
  sites_aov <- aov(intensity_mean ~ site * lubridate::year(date_start) * duration * distance, data = df)
return(sites_aov)
}

OISST_anov <- anova_func(df = OISST_upwell_base)
OISST_anov
CMC_anov <- anova_func(df = CMC_upwell_base)
CMC_anov

SACTN_anov <- sites_aov <- aov(intensity_mean ~ site * lubridate::year(date_start) * duration, data = SACTN_upwell_base)
SACTN_anov
```

@@

### 2019-09-24 Write_up/Paper1_writeup.Rmd
update

In order to create a time series of each of the remotely sensed SST data and also to compare whether the same upwelling patterns exist on the broad-scale against the *in situ* time series, shore-normal transects were extended from the *in situ* sites along the west coast of South Africa. Temperature values from the various satellite datasets were extracted at 15, 25, and 50 km along the shore-normal transect from the coast  (Fig. 1). The daily SST values obtained from the various remotely sensed datasets along each of the transects were used to produce an offshore time series that could accurately be used as a dataset to be compared against the *in situ* time series.

@@

In order to determine the change in upwelling signals at different distances from the coastline and to test which dataset showed clear, visible representation of these events, it was first necessary to define when upwelling was occurring. To do this it was necessary to determine a threshold value for identifying when upwelling was occurring. Given that upwelling is primarily caused by alongshore, equatorward winds, it was necessary to consider both temperature and wind values. To determine the upwelling condition, an upwelling index was calculated using the formula presented in Fielding and Davis (1989):

@@

In this equation, μ represents the wind speed (m/s) and θ represents the wind direction, which is measured in degrees. The coastal angle is the angle of each of the sites perpendicular to the coastline. The index relies heavily on wind speed and direction data in order to determine the intensity of upwelling likely to occur. Thereafter, the `exceedance()` function within the __`heatwaveR`__ package (RWS: Cite heatwaveR) was used to determine the consecutive number of days the upwelling index was at or above 1. In order to obtain a more approved threshold value, a percentile value was determined (RWS: Was it though? This sentence wouldn't make sense to someone that hadn't also seen the R code). To determine changes within the upwelling values, the `detect_event()` function within the __`heatwaveR`__ package was used. Thereafter, to determine the upwelling event metrics (frequency and intensity) the detect event function was used as a first filter with a threshold at or below the 25th percentile. Values obtained from the upwelling index were used as a second threshold. (RWS: The description of the statistical methodology here is unclear and needs to be re-written) An R script (https://github.com/AmierohAbrahams/Upwelling_MCS) was used to calculate the individual upwelling events for both the *in situ* and remotely sensed SST data. Because upwelling events were calculated relative to percentile exceedances, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occurred any time of the year with upwelling shown to be more dominant during spring and summer months. For the use of the heatwaveR package it was best to use a time series of 30 years (RWS: No. The 30 years comes from WMO recommendations and is not a requirement for the heatwaveR package specifically).

>>>>

In order to create a time series of each of the remotely sensed SST data and also to compare whether the same upwelling patterns exist on the broad-scale against the *in situ* time series, shore-normal transects were extended from the *in situ* sites along the west coast of South Africa. Temperature values from the various satellite datasets were extracted at 15, 25, and 50 km along the shore-normal transect from the coast shown as black boxes in Fig. 1. The daily SST values obtained from the various remotely sensed datasets along each of the transects were used to produce an offshore time series that could accurately be used as a dataset to be compared against the *in situ* time series. Offshore transects were used in order to generate a time series in this manner as they accurately represent the mesoscale temperatures we were interested in comparing against coastal upwelling. Had we only made use of the nearest remotely sensed pixels to compare against the SACTN time series this would simply only draw a comparison between datasets, and not the mesoscape acitivity within the ocean.

@@

In order to determine the change in upwelling signals at various different distances from the coastline, and to test which dataset showed a clear, and visible representation of these events, it was first necessary to define when upwelling was occurring; however to accomplish this it was first required to establish a threshold value for identifying when the phenomenon was in taking place. Given that upwelling is primarily caused by alongshore, equatorward winds, it was necessary to consider both temperature and wind values. These data values are instrumental in determining the specific environmental conditions and parameters required for upwelling to occur; to accomplish this an upwelling index was calculated using the formula presented in the work by Fielding and Davis (1989):

@@

In this equation, μ represents the wind speed (m/s) and θ represents the wind direction, which is measured in degrees, and the coastal angle as the angle of each of the sites perpendicular to the coastline. The index relies heavily on wind speed and direction data in order to determine the intensity of upwelling and the likelihood of the phenomenon occurring. The above equation produces a value called the upwelling index; for this study specifically, upwelling was classified to occur when the index is greater than 1. However to further ascertain a threshold for upwelling, the `exceedance()` function within the _`heatwaveR`_ package [@schlegel2017heat] was utilised, specifically to determine the consecutive number of days that the event must exceed for it to be regarded and classified as an upwelling event. However it should be noted that upwelling is known to vary on a seasonal basis and also may occur hourly. Therefore this study sought to establish the minimum duration for the analyses of upwelling as 1 day; the rationale behind this being that data from the SACTN dataset as well as the Satellite remotely sensed SST data is collected daily. Furthermore, it was also determined that a 25th percentile threshold for the upwelling index result lasting for >= 1 days matched with the upwelling index value of 1. As a result the 25th percentile threshold was used as a more acceptable value to predict upwelling events. In order to determine the upwelling metrics, the `detect_event()` function within the _`heatwaveR`_ package was used.An R script (https://github.com/AmierohAbrahams/Upwelling_MCS) was used to calculate the individual upwelling events for both the *in situ* and remotely sensed SST data. Because upwelling events were calculated relative to percentile exceedances, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occurred any time of the year with upwelling shown to be more dominant during spring and summer months. It must also be stated that due to irregular sampling efforts in the SACTN dataset the spatial and temporal distributions of upwelling is not necessarily even.
